### 2016-02-14 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
ESM

\end{center}
%\modulolinenumbers[1]
%\linenumbers

@@

%TG: Maybe useless, I'll add that just in the thesis.
%\section{Graphical abstract}

%\begin{figure}[!htbp]
%\centering
%    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
%\caption{Example of topological errors due to missing morphological data in living taxa.
%The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
%If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
%\label{Figure_missing_data_problem}
%\end{figure}

% NC: It's important to follow up on typo and grammar issues in the supp for things that were mentioned in the main text. For example, dropping https... MorphoBank...

@@

\subsection{Google Scholar (accessed 11th June 2015)}
To ensure we did not miss any extra matrices that were not available on one of these repositories, we ran a Google Scholar search on the 5th January % NC ??? It says 11th June above. 
using the following keywords:

@@

where \textit{order} was replaced by each of the taxonomic subdivision keywords listed above in turn.
For each taxonomic subdivision keyword we selected the 20 first papers published since 2010 resulting in 660 papers.
We selected only the 20 first results for each search term to avoid downloading large numbers of irrelevant articles, and because the rate of discovery of new matrices was very low and unlikely to be substantially improved by downloading more papers.
For example, in the 660 papers we downloaded, only 50 contained extra living OTUs and only contributed 425 OTUs to our total of 4950 OTUs (Figure ~\ref{Supp_figure_google_searches}).
We selected only articles published since 2010 because almost every recently published matrix contained some of the morphological characters and OTUs from previous studies, thus almost all older studies are represented in the matrices we collected.
For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil, marivaux2005anthropoid, seiffert2005basal, bloch2007new,bloch2007new, kay2008anatomy, silcox2008biogeographic, seiffert2009convergent,tabuce2009anthropoid, boyer2010astragalar, seiffert2010fossil, marivaux2013djebelemur, ni2013oldest}.

@@

\section{2 - Data collection reproducibility}
Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{github.com/TGuillerme/Missing_living_mammals}).
%Needs to fix this in GitHub
%1-Rarefaction ?
%2-Selecting the OTUs
%3-Analysing the availability and the data structure

>>>>

\end{center}

@@

% TO DO!
% NC: It's important to follow up on typo and grammar issues in the supp for things that were mentioned in the main text. For example, dropping https... MorphoBank...

@@

\subsection{Google Scholar (accessed 11th June 2015)}
To ensure we did not miss any extra matrices that were not available on one of these repositories, we ran a Google Scholar search on the 11th June using the following keywords:

@@

where \textit{order} was replaced by each of the taxonomic subdivision keywords listed above in turn.
For each taxonomic subdivision keyword we selected the 20 first papers published since 2010 resulting in 660 papers.
We selected only the 20 first results for each search term to avoid downloading large numbers of irrelevant articles, and because the rate of discovery of new matrices was very low and unlikely to be substantially improved by downloading more papers.
For example, in the 660 papers we downloaded, only 50 contained extra living OTUs and only contributed 425 OTUs to our total of 4950 OTUs (Figure ~\ref{Supp_figure_google_searches}).
We selected only articles published since 2010 because almost every recently published matrices contained some of the morphological characters and OTUs from previous studies, thus almost all older studies are represented in the matrices we collected.
For example, the six living primates used in \cite{ross1998phylogenetic} (\textit{Aotus trivirgatus, Galago demidoff, Lemur catta, Microcebus murinus, Nycticebus coucang and Saimiri sciureus}) and their associated characters are reused along with more living species and characters in \cite{seiffert2003fossil, marivaux2005anthropoid, seiffert2005basal, bloch2007new,bloch2007new, kay2008anatomy, silcox2008biogeographic, seiffert2009convergent,tabuce2009anthropoid, boyer2010astragalar, seiffert2010fossil, marivaux2013djebelemur, ni2013oldest}.

@@

\section{2 - Data collection reproducibility}
Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{github.com/TGuillerme/Missing_living_mammals}).

### 2016-02-14 Manuscript/BiologyLetters/Supplementary/ESM_2.tex
ESM

%TG: Maybe useless, I'll add that just in the thesis.
%\section{Graphical abstract}

%\begin{figure}[!htbp]
%\centering
%    \includegraphics[width=1\textwidth]{MissingDataFigure.pdf}
%\caption{Example of topological errors due to missing morphological data in living taxa.
%The phylogeny contains two clades, Aves and Mammalia, with molecular data (grey) for both but only morphological data (red) for Aves.
%If a mammalian fossil species (with no molecular data) is added to the phylogeny, it will erroneously branch with the Aves instead of the Mammalia because no morphological data will overlap between the fossil mammals and the living ones.}
%\label{Figure_missing_data_problem}
%\end{figure}

\section{Supplementary results - 1}
The following section contains a supplementary analysis looking at the sampling effort for mammalian orders at three different taxonomic levels (family, genus and species).
We fitted linear regressions %TG: does one calculate regressions? I would go for drawing regressions but that's not semantically correct... % NC: It's a model so you fit it
between the log of the number of sampled families/genera/species for each order against the log of the total number of families/genera/species in each order % NC: Taken from where? The taxonomy? Where do you get this total from?
and fixed the intercept at zero (because when no families/genera/species exist they cannot be sampled).
The slope of the regression gives an indication of the sampling effort: if the slope is 1, the number of sampled families/genera/species and the total number of families/genera/species are equal, i.e. all families/genera/species have been sampled. 
Slopes of less than 1 indicate lower sampling effort.
The code for this analysis is available here: \url{github.com/TGuillerme/Missing_living_mammals/blob/master/Analysis/Sampling_effort.R}.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_sampling_effort.pdf}
\caption{Sampling effort across mammalian orders at the three different taxonomic levels. The \textit{R$^2$} and the slope (\textit{b}, forced through the origin) are reported for each regression.} % NC: Should report adjusted r2 because sample size is small.
\label{Supp_Figure_sampling}
\end{figure}

% NC: Fix the y tick labels (use las = 1 to make them horizontal), what's up with panel c? the axes should be the same size as the others and why doesn't the x axis cover the range of the data? Also what are the three panels? explain in the legend and maybe label as a,b,c? Also axes labels should be log(sampled families) etc. 
%Supp mat is no excuse for sloppy figures.

This analysis suggests that sampling effort effectively reflects diversity for each taxonomic level, i.e. morphological characters were collected for each order according to its size (Figure \ref{Supp_Figure_sampling}).
However, it is worth noting that this relation is stronger at higher taxonomic levels (i.e. family-level, \textit{R$^2$}=0.99 \textit{b}=0.936) compared to the species-level (\textit{R$^2$}=0.852 \textit{b}=0.711).
These results might be simply due to the increasing number of groups at each taxonomic level (i.e. number of species $>$ genera $>$ families).
But it is also probable that mammalian morphological phylogeneticists have focused on resolving higher level relationships among mammalian orders (see Table 1 in the main text), resulting in better coverage at the family-level.

@@

% NC: I have copied this wording from the main text. Fix all the ones below please :)

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_CHIROPTERA.pdf}
\caption{Distribution of available morphological data across Chiroptera. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Chiroptera}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_CINGULATA.pdf}
\caption{Distribution of available morphological data across Cingulata. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Cingulata}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_DASYUROMORPHIA.pdf}
\caption{Distribution of available morphological data across Dasyuromorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Dasyuromorphia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_DIDELPHIMORPHIA.pdf}
\caption{Distribution of available morphological data across Didelphimorphia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Didelphimorphia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_DIPROTODONTIA.pdf}
\caption{Distribution of available morphological data across Diprotodontia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Diprotodontia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_ERINACEOMORPHA.pdf}
\caption{Distribution of available morphological data across Erinaceomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Erinaceomorpha}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_PILOSA.pdf}
\caption{Distribution of available morphological data across Pilosa. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Pilosa}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_CETARTIODACTYLA.pdf}
\caption{Distribution of available morphological data across Cetartiodactyla. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Primates}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_RODENTIA.pdf}
\caption{Distribution of available morphological data across Rodentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Rodentia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_SCANDENTIA.pdf}
\caption{Distribution of available morphological data across Scandentia. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Scandentia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_SORICOMORPHA.pdf}
\caption{Distribution of available morphological data across Soricomorpha. Edges are colored in grey when no morphological data is available or in blue when data is available.}
\label{Supp_Figure_Phylo-Soricomorpha}
\end{figure}

>>>>

\section{Supplementary results - 1}
The following section contains a supplementary analysis looking at the sampling effort for mammalian orders at three different taxonomic levels (family, genus and species).
We fitted linear regressions between the log of the number of sampled families/genera/species for each order against the log of the total number of families/genera/species in each order from \cite{bininda-emondsthe2007} and fixed the intercept at zero (because when no families/genera/species exist they cannot be sampled).
The slope of the regression gives an indication of the sampling effort: if the slope is 1, the number of sampled families/genera/species and the total number of families/genera/species are equal, i.e. all families/genera/species have been sampled. 
Slopes of less than 1 indicate lower sampling effort.
The code for this analysis is available here: \url{github.com/TGuillerme/Missing_living_mammals/blob/master/Analysis/Sampling_effort.R}.

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_sampling_effort.pdf}
\caption{Sampling effort across mammalian orders at the family (A), genus (B) and species (C) levels. The \textit{R$^2$} (adjusted) and the slope (\textit{b}, forced through the origin) are reported for each regression.}
\label{Supp_Figure_sampling}
\end{figure}

This analysis suggests that sampling effort effectively reflects diversity for each taxonomic level, i.e. morphological characters were collected for each order according to its size (Figure \ref{Supp_Figure_sampling}).
However, it is worth noting that this relation is stronger at higher taxonomic levels (i.e. family-level, adjusted \textit{R$^2$}=0.99 \textit{b}=0.936) compared to the species-level (\textit{R$^2$}=0.846 \textit{b}=0.71).
These results might be simply due to the increasing number of groups at each taxonomic level (i.e. number of species $>$ genera $>$ families).
But it is also probable that mammalian morphological phylogeneticists have focused on resolving higher level relationships among mammalian orders (see Table 1 in the main text), resulting in better coverage at the family-level.

\bibliographystyle{vancouver}
\bibliography{Supp_References}

@@

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_CHIROPTERA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Chiroptera. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Chiroptera}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_CINGULATA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Cingulata. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Cingulata}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_DASYUROMORPHIA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Dasyuromorphia. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Dasyuromorphia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_DIDELPHIMORPHIA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Didelphimorphia. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Didelphimorphia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_DIPROTODONTIA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Diprotodontia. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Diprotodontia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_ERINACEOMORPHA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Erinaceomorpha. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Erinaceomorpha}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_PILOSA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Pilosa. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Pilosa}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_CETARTIODACTYLA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Cetartiodactyla. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Primates}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_RODENTIA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Rodentia. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Rodentia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_SCANDENTIA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Scandentia. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Scandentia}
\end{figure}

\begin{figure}[!htbp]
\centering
    \includegraphics[width=1\textwidth]{Supp_figure_SORICOMORPHA.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters Soricomorpha. Blue branches indicate species with available coded anatomical characters}
\label{Supp_Figure_Phylo-Soricomorpha}
\end{figure}
